---
title: "SDS HOMEWORK 3"
author: "Jeremy Sapienza & Stefano D'Arrigo"
date: "21 Gennaio 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```

```{r include=FALSE}
load("./hw3_data.RData")
```

$$~$$

# Start with something...

$$~$$

After loading the data provided by the professor, create the covariance-matrix for each patient; Z-scale this matrix through the Fisher's Z-transform and after that create a unique matrix per group considering the olympic mean by $X_n =  (x_1, ... , x_{12})$ for each element in the matrix.

$$~$$

```{r}
# define a function in order to catch each person and return the correlation matrix between ROIs
get.cor.ROIs <- function(group, person){
  frame <- group[[person]]
  
  n   <- ncol(frame)
  nms <- names(frame)
  
  # takes efficiently the combinations, this is useful for large dataset!
  V1  <- rep(nms[1:(n-1)], seq(from=n-1, to = 1, by = -1)) 
  V2  <- unlist(lapply(1:(n-1), function(i)(nms[(i+1):n])))
  
  corr.ROIs <- data.frame(ROI_1=V1, ROI_2=V2) # create the correlation in which I will insert the values
  
  corr.ROIs$z.pearson.corr <- apply(corr.ROIs, 1, function(row) {
    atanh(cor(frame[row["ROI_1"]], frame[row["ROI_2"]])) # takes the sets of columns; apply the Z-transform to the
                                                          # correlation matrix
  })
  
  return(corr.ROIs)
}
```

$$~$$

Now, for each group create for each corresponding patient the matrix of correlations in the Z-scale.

$$~$$




```{r}
# create the matrix correlations for all patients
for(person in names(asd_sel)) assign(paste("z.trans.", person, sep=""), get.cor.ROIs(asd_sel, person))
for(person in names(td_sel)) assign(paste("z.trans.", person, sep=""), get.cor.ROIs(td_sel, person))
```

$$~$$

After that, we pooled the data considering the olympic mean between these datasets for each group, this is due to .... ecccccc

$$~$$

```{r}
# create a unique matrix for each group
unique.matrix <- function(group) {
  mtx <- z.trans.caltech_0051472[c("ROI_1", "ROI_2")] # create matrix with combinations
  
  mtx$olymp.mean <- apply(mtx, 1, function(row) {
  values <- c()
  for(person in names(group)) {
     frame <- get(paste("z.trans.", person, sep="")) # match the address of the string with the real variable
     elem <- frame[(frame[["ROI_1"]] == row["ROI_1"]) & (frame[["ROI_2"]] == row["ROI_2"]), "z.pearson.corr"] # select the correlation
     values <- c(values, elem)
  }
    #values <- values[values != min(values)]; values <- values[values != max(values)] # remove the min and max, we have continuous values, so the prob. to have twice the min or max is impossible..
    mean(values, trim = .2) # take the mean!
  })
  
  return(mtx)
}
```

$$~$$

Now, we call and create the matrix for each group of people as in this way:

$$~$$

```{r}
# call the creation of unique matrix
asd_sel.dt <- unique.matrix(asd_sel); head(asd_sel.dt)
td_sel.dt <- unique.matrix(td_sel); head(td_sel.dt)
```

$$~$$

# Association graphs


```{r include=FALSE}
# # pick through all correlation values the 80th percentile
# t.asd_sel <- apply(asd_sel.dt["olymp.mean"], 2, quantile, probs=0.8) 
# paste("the threshold at 80% percentile for the ASD subjects is: ", round(as.numeric(t.asd_sel), 3), sep = "")
# t.td_sel <- apply(td_sel.dt["olymp.mean"], 2, quantile, probs=0.8)
# paste("the threshold at 80% percentile for TD subjects is: ", round(as.numeric(t.td_sel), 3), sep = "")
```

Start to find the quantile value for these two groups. Since working on the Z-scale, the threshold is also in the Z-scale.

$$~$$

```{r}
z.t <- apply(rbind(asd_sel.dt["olymp.mean"], td_sel.dt["olymp.mean"]), 2, quantile, probs=0.8) 
paste("the threshold at 80% percentile for the subjects is: ", round(as.numeric(z.t), 3), sep = "")
```


$$~$$

Show what happens if we don't take into account the confidence interval for the Z-transformed $\rho$.

$$~$$

```{r}
# create the adj matrix 
nameVals <- sort(unique(unlist(asd_sel.dt[1:2]))) # set up storage matrix, get names for row and columns

# construct 0 matrix of correct dimensions with row and column names 
adj_mat <- matrix(0, length(nameVals), length(nameVals), dimnames = list(nameVals, nameVals)) 

# fill in the matrix with matrix indexing on row and column names 
adj_mat[as.matrix(asd_sel.dt[c("ROI_1", "ROI_2")])] <- 0

# set one if i achieve the decision rule 
for(idx in rownames(asd_sel.dt)){ 
  val <- abs(asd_sel.dt[idx, "olymp.mean"])
  if( val >= z.t ) {
    adj_mat[as.character(asd_sel.dt[idx, "ROI_1"]), as.character(asd_sel.dt[idx, "ROI_2"])] <- 1 
  } 
}

# To plot the correlation graph(s)
require(igraph, quietly = TRUE)

G1 <- graph_from_adjacency_matrix(adj_mat[1:20, 1:20], mode = "undirected") # show only 20 vertices
plot(G1, vertex.size = .5, vertex.color = rgb(0,0,1,.2), label.col = "black", curved = TRUE, main = "Marginal Correlation Graph", sub = "(No multiplicity adjustment)") 
```

$$~$$

Compute the confidence interval for $\bar{Z}_{12}(j,k)$. Remeber that we started computing the estimate of the covariance matrix for each person in each of the two groups, having:

$$
\Big\{\hat{\rho}^{(i)}_{j,k}\Big\}_{i=1}^{12}, \text{ having } j,k \in \{1,...,116\}, \text{ } j\neq k
$$
Then, we applied the Fisher's Z-transform:

$$
\Big\{\hat{Z}^{(i)}_{j,k}\Big\}_{i=1}^{12} = \Big\{atan\left(\hat{\rho}^{(i)}_{j,k}\right)\Big\}_{i=1}^{12}
$$

Finally, we computed the estimator:

$$
\overline{Z}_{12}(j,k) = \frac{1}{12}\sum_{i=1}^{12}\hat{Z}^{(i)}_{j,k} \text{ ,} 
$$
remembering that $\hat{Z}^{(i)}_{j,k}$ are $IID$, being function of $\hat{\rho}^{(i)}_{j,k}$, respectively.

Now, we calculate the confidence interval for each $\overline{Z}_{12}(j,k)$. Starting from:

$$
\frac{\overline{Z}_{12}(j,k) - z_{j,k}}{\sigma_{j,k} / \sqrt{12}}, \\ 
\text{ where } \sigma_{j,k} = \frac{1}{\sqrt{145 - 3}}
$$

and applying the Bonferroni's correction:

$$
\frac{\alpha}{m}, \text{ where } m = \begin{pmatrix} D \\ 2 \end{pmatrix}, \text{ } D=116 
$$

we end up with:

$$
C_{12}^{Z(j,k)}\Big(\frac{\alpha}{m}\Big) = \bigg[\overline{Z}_{12}(j,k) \mp z_{\alpha/2m} \cdot \frac{\sigma_{j,k}}{\sqrt{12}}\bigg]
$$

$$~$$

```{r}
conf.int <- function(dt) {
  # Asymptotic variance
  n.samp <- 145 # number of IID samples on which we computed the covariance matrix
  sigma <- 1 / sqrt(n.samp - 3) # standard deviation of the (j,k)-th Z-transformed Pearson coefficent 
  n.p <- 12 # number of people in each of the two groups
  se <- sigma / sqrt(n.p) # standard error of the estimator Z12
  
  # Confidence interval - Setup
  alpha <- 0.05 # confidence level WITHOUT the Bonferroni's correction
  m <- (116 * 115) / 2 # number of hypothesis
  
  # Confidence interval
  cint <- sapply(dt$olymp.mean, function(z.i) {
    list(c(lb = z.i - se * qnorm(1 - alpha / (2*m)), 
           ub = z.i + se * qnorm(1 - alpha / (2*m))))
  })
  return(cint)
}
```

Finally:

```{r}
## Compute the confidence interval
asd_sel.dt$cint <- conf.int(asd_sel.dt)
td_sel.dt$cint <- conf.int(td_sel.dt)
```








