---
title: "SDS HOMEWORK 3"
author: "Jeremy Sapienza & Stefano D'Arrigo"
date: "21 Gennaio 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```

```{r include=FALSE}
load("D:/UNIVERSITA'/MASTER DEGREE - La Sapienza/SDS - Statistical Methods in Data Science/HOMEWORK/HW3/sapienza_darrigo/hw3_data.RData")
```

$$~$$

# Start with something...

$$~$$

After loading the data provided by the professor, create the covariance-matrix for each patient and after that create a unique matrix per group considering the olympic mean by $X_n =  (x_1, ... , x_{12})$ for each element in the matrix.

$$~$$

```{r}
# define a function in order to catch each person and return the correlation matrix between ROIs
get.cor.ROIs <- function(group, person){
  frame <- group[[person]]
  
  n   <- ncol(frame)
  nms <- names(frame)
  
  # takes efficiently the combinations, this is useful for large dataset!
  V1  <- rep(nms[1:(n-1)], seq(from=n-1, to = 1, by = -1)) 
  V2  <- unlist(lapply(1:(n-1), function(i)(nms[(i+1):n])))
  
  corr.ROIs <- data.frame(ROI_1=V1, ROI_2=V2) # create the correlation that I will insert the values
  
  corr.ROIs$pearson.corr <- apply(corr.ROIs, 1, function(row) {
    cor(frame[row["ROI_1"]], frame[row["ROI_2"]]) # takes the sets of columns
  })
  
  return(corr.ROIs)
}
```

$$~$$

Now, for for each group create for each corresponding patient the matrix of correlations

$$~$$

```{r}
# create the matrix correlations for all patients
for(person in names(asd_sel)) assign(paste("corr.", person, sep=""), get.cor.ROIs(asd_sel, person))
for(person in names(td_sel)) assign(paste("corr.", person, sep=""), get.cor.ROIs(td_sel, person))
```

$$~$$

After that, we pooled the data considering the olympic mean between these datasets for each group, this is due to .... ecccccc

$$~$$

```{r}
# create a unique matrix for each group
unique.matrix <- function(group) {
  mtx <- corr.caltech_0051472[c("ROI_1", "ROI_2")] # create matrix with combinations
  
  mtx$olymp.mean <- apply(mtx, 1, function(row) {
  values <- c()
  for(person in names(group)) {
     frame <- get(paste("corr.", person, sep="")) # match the address of the string with the real variable
     elem <- frame[(frame[["ROI_1"]] == row["ROI_1"]) & (frame[["ROI_2"]] == row["ROI_2"]), "pearson.corr"] # select the correlation
     values <- c(values, elem)
  }
    values <- values[values != min(values)]; values <- values[values != max(values)] # remove the min and max, we have continuous values, so the prob. to have twice the min or max is impossible..
    mean(values) # take the mean!
  })
  
  return(mtx)
}
```

$$~$$

Now, we call and create the matrix for each group of people as in this way:

$$~$$

```{r}
# call the creation of unique matrix
asd_sel.dt <- unique.matrix(asd_sel); head(asd_sel.dt)
td_sel.dt <- unique.matrix(td_sel); head(td_sel.dt)
```



