---
title: "SDS HOMEWORK 3"
author: "Jeremy Sapienza & Stefano D'Arrigo"
date: "21 Gennaio 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```

```{r include=FALSE}
load("./hw3_data.RData")
```

$$~$$

# Start with something...

$$~$$

After loading the data provided by the professor, create the covariance-matrix for each patient and after that create a unique matrix per group considering the olympic mean by $X_n =  (x_1, ... , x_{12})$ for each element in the matrix.

$$~$$

```{r}
# define a function in order to catch each person and return the correlation matrix between ROIs
get.cor.ROIs <- function(group, person){
  frame <- group[[person]]
  
  n   <- ncol(frame)
  nms <- names(frame)
  
  # takes efficiently the combinations, this is useful for large dataset!
  V1  <- rep(nms[1:(n-1)], seq(from=n-1, to = 1, by = -1)) 
  V2  <- unlist(lapply(1:(n-1), function(i)(nms[(i+1):n])))
  
  corr.ROIs <- data.frame(ROI_1=V1, ROI_2=V2) # create the correlation in which I will insert the values
  
  corr.ROIs$pearson.corr <- apply(corr.ROIs, 1, function(row) {
    cor(frame[row["ROI_1"]], frame[row["ROI_2"]]) # takes the sets of columns
  })
  
  return(corr.ROIs)
}
```

$$~$$

Now, for for each group create for each corresponding patient the matrix of correlations

$$~$$

```{r}
z_transform <- function(cor.mat) {
  return(apply(cor.mat, 1, function(row){
    atanh(as.numeric(row["pearson.corr"])) 
  }))
}
```




```{r}
# create the matrix correlations for all patients
for(person in names(asd_sel)) assign(paste("corr.", person, sep=""), z_transform(get.cor.ROIs(asd_sel, person)))
for(person in names(td_sel)) assign(paste("corr.", person, sep=""), z_transform(get.cor.ROIs(td_sel, person)))
```

$$~$$

After that, we pooled the data considering the olympic mean between these datasets for each group, this is due to .... ecccccc

$$~$$

```{r}
# create a unique matrix for each group
unique.matrix <- function(group) {
  mtx <- corr.caltech_0051472[c("ROI_1", "ROI_2")] # create matrix with combinations
  
  mtx$olymp.mean <- apply(mtx, 1, function(row) {
  values <- c()
  for(person in names(group)) {
     frame <- get(paste("corr.", person, sep="")) # match the address of the string with the real variable
     elem <- frame[(frame[["ROI_1"]] == row["ROI_1"]) & (frame[["ROI_2"]] == row["ROI_2"]), "pearson.corr"] # select the correlation
     values <- c(values, elem)
  }
    #values <- values[values != min(values)]; values <- values[values != max(values)] # remove the min and max, we have continuous values, so the prob. to have twice the min or max is impossible..
    mean(values, trim = .2) # take the mean!
  })
  
  return(mtx)
}
```

$$~$$

Now, we call and create the matrix for each group of people as in this way:

$$~$$

```{r}
# call the creation of unique matrix
asd_sel.dt <- unique.matrix(asd_sel); head(asd_sel.dt)
td_sel.dt <- unique.matrix(td_sel); head(td_sel.dt)
```

# Association graphs

Start to find the quantile value for these two groups.

# ```{r}
# # pick through all correlation values the 80th percentile
# t.asd_sel <- apply(asd_sel.dt["olymp.mean"], 2, quantile, probs=0.8) 
# paste("the threshold at 80% percentile for the ASD subjects is: ", round(as.numeric(t.asd_sel), 3), sep = "")
# t.td_sel <- apply(td_sel.dt["olymp.mean"], 2, quantile, probs=0.8)
# paste("the threshold at 80% percentile for TD subjects is: ", round(as.numeric(t.td_sel), 3), sep = "")
# ```

```{r}
t <- apply(rbind(asd_sel.dt["olymp.mean"], td_sel.dt["olymp.mean"]), 2, quantile, probs=0.8) 
paste("the threshold at 80% percentile for the subjects is: ", round(as.numeric(t), 3), sep = "")
```


```{r}
# create the adj matrix 
nameVals <- sort(unique(unlist(asd_sel.dt[1:2]))) # set up storage matrix, get names for row and columns

# construct 0 matrix of correct dimensions with row and column names 
adj_mat <- matrix(0, length(nameVals), length(nameVals), dimnames = list(nameVals, nameVals)) 

# fill in the matrix with matrix indexing on row and column names 
adj_mat[as.matrix(asd_sel.dt[c("ROI_1", "ROI_2")])] <- 0

# set one if i achieve the decision rule 
for(idx in rownames(asd_sel.dt)){ 
  val <- abs(asd_sel.dt[idx, "olymp.mean"])
  if( val >= t ) {
    adj_mat[as.character(asd_sel.dt[idx, "ROI_1"]), as.character(asd_sel.dt[idx, "ROI_2"])] <- 1 
  } 
}

# To plot the correlation graph(s)
require(igraph, quietly = TRUE)

G1 <- graph_from_adjacency_matrix(adj_mat[1:20, 1:20], mode = "undirected") 
plot(G1, vertex.size = .5, vertex.color = rgb(0,0,1,.2), label.col = "black", curved = TRUE, main = "Marginal Correlation Graph", sub = "(No multiplicity adjustment)") 
```

$$~$$


```{r}
# Fisher Z-transform
z <- sapply(asd_sel.dt$olymp.mean, atanh)

# Asymptotic variance
n <- 12
sigma <- 1 / sqrt(n - 3)

# Confidence interval
alpha <- 0.05 # level
asd_sel.dt$cint <- sapply(z, function(z.i) {
  list(tanh(c(lb = z.i - sigma * qnorm(1 - alpha / 2), 
         ub = z.i + sigma * qnorm(1 - alpha / 2))))
})
```







